---
title: "Numeric Data - How To Build 8 Individual Models.qmd"
format: html
---

# How to build numeric models that beat Kaggle data science competition winners

Welcome to the chapter on building individual numeric models. This is the first of two chapters to analyze numeric data, the second (next) chapter will build on the foundation of this chapter and show how to build ensembles of numeric models. In many ways, this is an excellent place to start. Everyone is familiar with numeric data, and this pair of chapters will show you how to get the most accurate results from virtually any numeric data set.

Once you are comfortable analyzing numeric data, you will be able to work with data that predicts all of these types of results and many more:

-   Sale price of houses

-   Mileage (miles per quantity of fuel) for vehicles

-   Food quality ratings (if they are numeric)

-   Student performance

-   Energy Efficiency

-   Strength of Concrete

-   Parkinson's Telemonitoring

-   Soybean yield

Many other types of data will also work with the methods to solve numeric data sets we will learn in these two chapters. You will be equipped to do an enormous amount of good with the skills you learn in these two chapters analyzing numeric data sets. This promises to be a lot of fun and very practical, so let's get started!

------------------------------------------------------------------------

## Summary of what is coming

This chapter will show how to build individual numeric models of housing prices, including the following components:

-   A Kaggle data science competition to use as a baseline. Our models will beat the best score in the competition.

-   How to read in a data set

-   How to set up the data set so the analysis are much easier to perform

-   How to randomize the data before you analyze it so the results generalize more accurately

-   How to set up the data to be randomly resampled

-   How to set up optimized (tuned) models

-   How to set up summary plots

-   How to set up summary reports

-   How to build an intermediate level solution (uses numeric data and strings)

-   How to build an advanced solution (working with numeric data, strings, factors, and a stratified column to model prices of BMW cars)

-   Actual code you may use or adapt for your own projects and keep in your code library for future use

------------------------------------------------------------------------

## A Kaggle data science competition to use as a baseline for the first example

Our very first example is one I am looking forward to sharing with you, but not for all the reasons you might expect. We will be looking at the famous Boston housing data set. The data set was originally created in 1970 and used in an article published in 1978. Compared to almost anything else you will ever work with in your entire career, this data is ancient!

One huge advantage of using the Boston Housing data set is that it has been used by a large number of previous investigators, so it serves as a benchmark for the accuracy of algorithms.

The data also has a substantial number of problems which we are going to work to identify, and that's one of the reasons I'm interested in sharing it with you. The data set has been criticized, correctly, for several issues. As we build our library of solutions, we want to test if these issues show up. Assuming they show up for the Boston housing data set (and the issues will show up!), it is a reasonable conclusion that similar issues will also show up for other data sets, and in my experience, this is very true.

**Better than AI Solution:** Every AI system is trained on lots of data, and designed to predict the most likely outcome. Our solutions will return great results (of course), but we will use human intelligence to see things no AI system is going to return to you. One of those will be training you to use your judgement when there are problems in the data. The problems can be how the data was recorded, poor ethical choices by prior researchers, clear examples of discrimination, and much more.

Once you learn how to see these types of problems, it will be easier to find them in other data sets.

One of the areas that can set you apart as a data scientist is being able to identify these kinds of issues, and the Boston housing data set is a great example of this exact type of solution. Once you see how this works, feel free to use it on other data sets and include those results in your portfolio.

::: {.callout-note appearance="minimal"}
"In a world of AI, your freedom is in your unpredictability. You have a whole host of large language models all of which are trained to do the most predictable thing next. One of the things that differentiates humans from AI (amongst many) is that we're not the most efficient of beings. The things that are the most meaningful to us are not necessarily the things that are the most predictable to us. "

"As a harmonist, it was one of my greatest joys to take a song or a melody and find kind of the most unexpected things to do with it. It would be like whatever you think I'm gong to do, I'm going to do the opposite of that"

" The right amount of irreverence can be so liberating."\
\
Jacob Collier, <https://youtube.com/shorts/QnGXDfi_itI?si=NM2ispZEpwcUuNUX>
:::

::: callout-note
"AI will not solve poverty, because the conditions that lead to societies that pursue profit over people are not technical. AI will not solve discrimination... the cultural patterns... are not technical."

— Joy Buolamwini (Computer Scientist & Digital Activist), *Unmasking AI: My Mission to Protect What Is Human in a World of Machines*, 2023, p. xix
:::

------------------------------------------------------------------------

**First** Let's start by becoming familiar with the Boston housing data set.

You may either download the data from the Kaggle.com website, or load it from the MASS package. To get the Boston Housing data set fro the MASS package, and look at the first ten rows:

```{r Get the Boston Housing data set from the MASS package}

df <- MASS::Boston
head(df, n = 10)
```

**Data dictionary**:

This will look at each feature and provide the information from the original article:

CRIM per capita crime rate by town

ZN proportion of residential land zoned for lots over 25,000 sq.ft.

INDUS proportion of non-retail business acres per town

CHAS Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)

NOX nitric oxides concentration (parts per 10 million)

RM average number of rooms per dwelling

AGE proportion of owner-occupied units built prior to 1940

DIS weighted distances to five Boston employment centres

RAD index of accessibility to radial highways

TAX full-value property-tax rate per \$10,000

PTRATIO pupil-teacher ratio by town

B 1000(Bk - 0.63)\^2 where Bk is the proportion of blacks by town

LSTAT % lower status of the population

MEDV Median value of owner-occupied homes in \$1000's \[\^libstat\]

**Data summary: Numeric Summary**

Let's take a look at the summary of each of the 14 columns, and see what we can learn:

```{r read in the Boston Housing data set, set up a summary}

df <- MASS::Boston
summary(df)
```

The summary provides six measures for each feature:

-   Minimum

-   1st quartile (25th percentile)

-   Median

-   Mean

-   3rd quartile (75th percentile)

-   Maximim

::: callout-note
Another advantage you have over AI is you can be surprised. Looking for surprises in the data, or data summaries, can be a great way to have a more accurate understanding of the data or the data summary that AI cannot match.
:::

One of the immediate observations in the summary is the features have a wide range of values. For example, crim ranges from 0.00632 to 88.97620, but rm ranges from 3.561 to 8.780.

Next, look for results that are surprising. A large surprise is the maximum housing price (measured in thousands of dollars): 50.00. What seems odd is the maximum price is an exact round number. The analysis we will conduct here will demonstrate the medv values were truncated at 50.00, and very likely not the true house value for the properties at 50.00.

::: {.callout-note appearance="minimal"}
Another advantage you have over AI is humans want to compare and compete. An AI system will only do that if it is asked to do so. If we look at the Boston Housing data set, not only are there a large number of previous results we can use for comparison, there are also *corrections to the original data set* which have been published. Belsley, Kuh & Welsch published a corrected version of the original Boston Housing data set in their book, Regression Diagnostics: Identifying Influential Data and Sources of Collinearity is an excellent example, and the data set is available at: <https://lib.stat.cmu.edu/datasets/boston_corrected.txt> \[\^Belsey\]

This is a very good practice: In addition to looking for the results from other investigators, search for corrections to the original data.
:::

------------------------------------------------------------------------

**Data summary: Visual**

It is always possible to look at data in multiple ways. For our example, we will look at the Boston Housing data set in a visual way. Here is a code chunk that provides boxplots for all the numeric values in a data set:

#### Boxplots:

```{r Boston Housing boxplots of the numeric data}

library(tidyverse)
df %>%
  tidyr::gather(key = "var", value = "value") %>%
  ggplot2::ggplot(aes(x = "", y = value)) +
  ggplot2::geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  ggplot2::facet_wrap(~var, scales = "free") +
  ggplot2::labs(title = "Boxplots of the numeric data")
```

We can see a lot from the box plots. A few examples:

• Several of the features have a large inter-quartile range, including age, tax, and rad.

• Several of the features have a very small inter-quartile range, including black, chas and crim.

•Several of the features have a large number of outliers, such as black, crim, lstat, medv. rm and zn.

• Several of the features have few (if any) outliers, such as age, indus, nox, rad and tax.

#### Histograms

```{r Histograms of each numeric column}

ggplot2::ggplot(tidyr::gather(df, cols, value), aes(x = value)) +
  ggplot2::geom_histogram(bins = round(nrow(df))) +
  ggplot2::facet_wrap(. ~ cols, scales = "free") +
  ggplot2::labs(title = "Histograms of each numeric column.")
```

The histograms tell us a lot of information about the data. For example, if we look at the histogram for medv, we can see the house values are truncated at \$50.00, and there are far more of those than any other value in the histogram.

#### Target variable vs each of the features

```{r Scatterlots of target vs each feature}

df %>%
  tidyr::gather(-medv, key = "var", value = "value") %>%
  ggplot2::ggplot(aes(x = value, y = medv)) +
  ggplot2::geom_point() +
  ggplot2::facet_wrap(~ var, scales = "free") +
  ggplot2::theme_bw()+
  ggplot2::labs(title = "y (predictor variable) vs target variables")
```

This set of plots gives us an excellent summary of the data. It plots the target value (medv) against each of the features, in alphabetical order. The x-axis is the range for each feature (such as age or tax) and the y-values are medv.

We can see (again) the medv values are truncated at 50, as the top value on each of the scatterplots.

------------------------------------------------------------------------

**The goal** in this case is to model the last column, medv, or the median value of the property, measured in thousands of dollars. For example, if a property has an MEDV value of 28.7, that property is valued at \$28,700.

Another issue that will help you become a better data scientist is to recognize the accuracy of labels of columns. In this particular example, our goal is to model medv. That column is listed as the median value of the house price, but that is wrong. It is not the median value of the house price, medv IS the value of the house price.

A very good opportunity with the Boston housing data set is that it has been used many times in Kaggle data science competitions. Therefore, we will have a great opportunity to compare our results to the results of several other data science competitions. You will see in a moment how to find even more results if you wish, but the results you will see from other Kaggle data science competitions on the status that are very typical.

We will start with a typical Kaggle data science competition. \[\^Kaggle1\] This competition has:

• 49 Entrants

• 49 Teams

• 699 Submissions

The most accurate results in this competition reported the following results on the public leaderboard (top six rows only shown):

| Team     | Lowest RMSE | Number of Submittals |
|----------|-------------|----------------------|
| 20204184 | 2.09684     | 75                   |
| 20181838 | 2.11690     | 30                   |
| 20181850 | 2.11691     | 31                   |
| 20201792 | 2.11692     | 16                   |
| 20181851 | 2.13075     | 41                   |
| 20181782 | 2.20019     | 8                    |

We will come up with a very easy solution that will beat all these results, runs very fast and without any errors, and that you can save in your code library to use for other numeric data set analyses, thus saving you the time to re-create a numeric data set solution when you already have it ready.

## How to read in a data set

Now that we know our goal is to minimize the root mean squared error of the home value, our first step is to read in the data set.

The Boston Housing data set is available in several places, we will read it in from the MASS package, as follows:

```{r Reading in the Boston Housing data set}
library(MASS)
df <- MASS::Boston
```

Let's take a quick look at our data set:

```{r First look at our data set}
head(df, n= 10)
```

We can see ten columns, and all of the columns are numeric. Later on we will work with data sets where some columns are not numeric.

## How to set up the data set so the analysis is much easier to perform in the future

We're going to start at the end and work backwards here. The target column in this data set is medv. However, the name of the target column for other data sets will be different. We are going to take steps to set up all our data set in a similar way. This will make it much easier to work with a very wide range of data sets.

You might have noticed we had to include 'medv' in our previous visualizations. Feel free to re-read the code examples if there are any questions about the name of the target variable. We can eliminate that step moving forward by always changing the name of the target to 'y' as soon as possible, in every data set we use.

> We can save an enormous amount of time by standardizing the data set when we read it in.

Let's set this up to work with virtually ANY data set, not just Boston Housing. We'll do that in two steps:

1.  Rename the target column as 'y'

2.  Always move the target column to the right most position

These two steps will make it much easier for our modeling system to work with any data set we read it, and allow us to use a very wide range of modeling methods, as you will see.

::: {.callout-note appearance="simple"}
The big secret at this point is to format all the data sets in the same layout. Not just the Boston Housing data set, but all the other data sets we will use, too. The last column will always be the target column, and always named 'y'. These simple changes will speed up the coding and analysis process when we use a range of modeling methods, as you will see in just a moment.

This is another example of starting at the end and working backwards. We know what we want in the end, a data set that can be used by all the modeling methods, so let's create it up front at the beginning.
:::

```{r Move the target column to the last position, rename the target column as y}
library(tidyverse)
df <- df %>% relocate(medv, .after = last_col())
df <- df %>% dplyr::rename(y = medv)
head(df)
```

## How to always randomize the data before you analyze it

This is a single line of code, but can make your final analysis much stronger than if it is not used. We are going to randomize the data, so the result is much more like the real world than if we did not randomize the data. This code randomizes all the rows, and keeps the name of the data (as df) and the last column (as y). You can see the data has been randomized because the row numbers are no longer in sequence.

```{r Randomize all the rows of the data set}
df <- df[sample(nrow(df)), ]
head(df)
```

This is another simple step that can great improve the reproducibility of your results. Many data sets come with the data already sorted by one or more of the columns. Therefore, the learner will learn where the data is rather than the actual underlying structure, and those results will not transfer to new data. It greatly improves the generalizability of your results if you sort the data before you do the analysis.

<h2>How to test if the data is randomized</h2>

The way to test if your data is randomized is to look at the row names. The Boston housing data set does not have actual role numbers, but it does have wrong names, and this is common in data science. You should see that the row names are not in sequence. The row names are not a column in the state set, which makes it another good example to learn from. If you want to see the row names, this code will reveal the row names to you.

```{r show row names}
row.names(df)
str(row.names(df))
```

You do not need to do anything in this case with the row names, but you might need to do that in the future with other data sets.

As another quick example of row names. The first column in the mtcars data set is not a normal column, so it cannot be called by the normal method, such as mtcars\$mpg. The very first column of mtcars can be retrieved the same way we just retrieved the road names for the Boston housing data set, by using the row.names function:

```{r How to get the row names from MTCars}

row.names(mtcars)
```

## How to split data into train and test sets

We'll use 60% train and 40% test, but those numbers are not set in stone. You may use values that work best for you.

We will start with the entire randomized data set. We will set the first 60% of the rows as the training set, and the remaining 40% as the test set.

We will be fitting our models on the training set, and making predictions on the test (holdout) set in the next section of this chapter. We will also take a quick look at the train and test sets.

The sample function can improve the results you achieve. This will randomly split the data into a number of groups. We will split the data into a train set (60% of the data) and a test set (40% of the data).

Let's take a look at the results so far, and test if we have everything we need:

```{r How to split data into train and test sets}
sample <- sample(c(TRUE, FALSE), nrow(df), replace=TRUE, prob=c(0.6,0.4))
train  <- df[sample, ]
test   <- df[!sample, ]
head(train)
head(test)
```

Test if we are ready to set up our models (and this will work with a wide range of numeric data sets):

Rows are randomized

Our target column is named y

Our target column is the right-most column

Looks good so far, everything passed the tests up to this point.

> I suggest saving everything up to this point in your data science library, you will be using it often in your career!

------------------------------------------------------------------------

## The data is set up, let's start building models!

#### How to set up optimized (tuned) formulas

We are going to start with numeric eight modeling methods. Since our data is already set up to be easy to use, all our modeling methods will be virtually identical:

model(y \~ ., data = train)

Here is how to install our first eight numeric modeling methods (you will only need to do this once):

```{r Introducing our first eight numeric modeling methods}

# we will install the packages if needed:

install.packages('ipred', repos = "http://cran.us.r-project.org") # for the bagging modeling method
install.packages('arm', repos = "http://cran.us.r-project.org") # for the bayesglm modeling method
install.packages('brnn', repos = "http://cran.us.r-project.org") # for the bayesrnn modeling method
install.packages('earth', repos = "http://cran.us.r-project.org") # for the earth modeling method
install.packages('e1071', repos = "http://cran.us.r-project.org") # for tuned models, starting with random forest and support vector machines
install.packages('tree', repos = "http://cran.us.r-project.org") # for the tree modeling method
```

Now we are ready to build our first eight models.

We will always be fitting the models on the training data, then making predictions and checking accuracy on the testing (holdout) data using the trained models.

```{r Building models}
# Begin by loading the modeling packages we will be using:
library(arm)
library(brnn)
library(earth)
library(e1071)
library(ipred)
library(randomForest)
library(tree)

# Now we will build each of the eight numeric models as applied to the Boston Housing data set

bagging <- ipred::bagging(y ~ ., data = train)
bayesglm <- arm::bayesglm(y ~ ., data = train, family = gaussian(link = "identity"))
bayesRNN <- brnn::brnn(x = as.matrix(train), y = train$y)
earth <- earth::earth(y ~ ., data = train)
linear <- lm(y ~ ., data = train)
randForest <- e1071::tune.randomForest(y ~ ., data = train)
svm <- e1071::tune.svm(y ~ ., data = train)
tree <- tree::tree(y ~ ., data = train)
```

### Measure the root mean squared error for each of our eight models

Now that we have the eight models built, let's check the root mean squared error on the testing data. These will all be set up in a very similar manner:

model_rmse \<- Metrics::rmse(actual = test\$y, predicted = predict(object = trained model, newdata = test)

All we need to do now is substitute the name of each of the eight models, and check the results:

```{r Measure RMSE for each of our five models}
bagging_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = bagging, newdata = test))
bayesglm_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = bayesglm, newdata = test))
bayesRNN_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = bayesRNN, newdata = test))
earth_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = earth, newdata = test))
linear_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = linear, newdata = test))
randForest_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = randForest$best.model, newdata = test))
svm_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = svm$best.model, newdata = test))
tree_rmse <- Metrics::rmse(actual = test$y, predicted = predict(object = tree, newdata = test))
```

Let's look at the results, sorted from most accurate to least accurate:

```{r Look at the results of our five models}
results <- data.frame('Name' = c('bagging', 'bayesGLM', 'bayesRNN', 'Earth', 'Linear', 'RandomForest', 'SVM', 'Treet'), 'RMSE' = c(bagging_rmse, bayesglm_rmse, bayesRNN_rmse, earth_rmse, linear_rmse, randForest_rmse, svm_rmse, tree_rmse))
results <- results %>% arrange(RMSE)
results
```

These results look good! Our best result beat all the other results in our analysis by an enormous amount, and also beat all the 699 results from the Kaggle.com competition we looked at earlier in the chapter. It only took a few lines of code, and this is code you can re-use as often as you need, to analyze other numeric data sets.

> I very strongly recommend saving these code snippets in your personal code library, so you have easy access to them and can use them over and over when you are analyzing numeric data sets. They work with the Boston Housing data set, and will work with virtually any other tidy data set.

As a teaser, when we start making ensembles of numeric data in the next chapter, our lowest root mean squared error we will beat our current best result by a very nice margin! We will see that in the next chapter.

> A very good way to improve your work and your code is to ask **'What is one thing I can do to improve the results?'** If you look for any way to improve your results, that will be too vague, and the human brain does not solve problems that way. If you ask, "What can I do?", again that is too vague. Give your brain a very specific goal, and it will work to solve it.

One thing we can do to improve our results is to randomly resample the results multiple times. What we will do is randomize the data, run the analysis and take the mean of the results. That will be repeated 25 times, and the mean of those 25 resamples will be a more accurate representation of the data than doing it just once.

There is a fairly fast law of diminishing returns on resampling. I've tested and found that approximately 25 resamples gives results that are not significantly different than doing 100 or 1,000 resamples. I recommend always testing everything to be sure.

## Now let's run the same code, but randomly resample it 25 times.

Our only difference will be tracking the mean of each root mean squared error (rmse) value, and taking the mean at the end.

Let's do it! This will take 1-2 minutes to run on a normal computer.

```{r randomly resample 25 times}

# Initialize the values to 0
bagging_predicted <- 0
bagging_rmse <- 0
bayesrnn_rmse <- 0
earth_rmse <- 0
linear_rmse <- 0
randForest_rmse <- 0
svm_rmse <- 0
tree_rmse <- 0

for (i in 1:25) {
  # Start by randomly resampling the data
  sample <- sample(c(TRUE, FALSE), nrow(df), replace=TRUE, prob=c(0.6,0.4))
train  <- df[sample, ]
test   <- df[!sample, ]

# Build the models on the current data set
bagging <- ipred::bagging(y ~ ., data = train)
bayesrnn <- brnn::brnn(x = as.matrix(train), y = train$y)
earth <- earth::earth(y ~ ., data = train)
linear <- lm(y ~ ., data = train)
randForest <- e1071::tune.randomForest(y ~ ., data = train)
svm <- e1071::tune.svm(y ~ ., data = train)
tree <- tree::tree(y ~ ., data = train)

# Make predictions and measure root mean square accuracy for each resample.
bagging_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = bagging, newdata = test))
bagging_predicted[i] <- predict(object = bagging, newdata = test)

bayesrnn_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = bayesrnn, newdata = test))
earth_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = earth, newdata = test))
linear_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = linear, newdata = test))
randForest_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = randForest$best.model, newdata = test))
svm_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = svm$best.model, newdata = test))
tree_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = tree, newdata = test))

# summarize the results
results <- data.frame('Name' = c('bagging', 'bayesRNN', 'Earth', 'Linear', 'RandomForest', 'SVM', 'Treet'), 'RMSE' = c(mean(bagging_rmse), mean(bayesrnn_rmse), mean(earth_rmse), mean(linear_rmse), mean(randForest_rmse), mean(svm_rmse), mean(tree_rmse)))
results <- results %>% arrange(RMSE)
}
results
```

<h2>Add a 95% confidence interval</h2>

Now that we've seen how to resample data, let's add a 95% confidence interval to the root means squared error results.

R provides four functions that work with different measurements. From the description (stats::TDist):

`dt` gives the density, `pt` gives the distribution function, `qt` gives the quantile function, and `rt` generates random deviates.

We will be using qt here.

A confidence interval is "a range of values likely to contain a population parameter with a certain level of confidence." \[2CI\]

A confidence interval is defined as:

(point estimate) ± (critical value) \* (standard error)

We will use a 95% confidence interval for our results. The format will be:

modeling_margin \<- stats::qt(0.975, df = nrow(test)\*bagging_holdout_RMSE_mean/sqrt(nrow(test)).

For example, here is the margin for the bagging modeling method:

```{r Calculating a 95% margin on the results using the bagging model}

  bagging_margin <- stats::qt(0.975, df = (nrow(test))*mean(bagging_rmse)/sqrt(nrow(test)))
  bagging_lower95 <- mean(bagging_rmse) - bagging_margin
  bagging_upper95 <- mean(bagging_rmse) + bagging_margin
  
  bagging_lower95
  bagging_upper95
```

Now that it's working for the bagging modeling method, we can easily set it up for the other modeling methods:

```{r Setting up confidence intervals for the remaining modeling methods}

bayesrnn_margin <- stats::qt(0.975, df = nrow(test)-1)*sd(bayesrnn_rmse)/sqrt(nrow(test))
bayesrnn_lower95 <- mean(bayesrnn_rmse) - bayesrnn_margin
bayesrnn_upper95 <- mean(bayesrnn_rmse) + bayesrnn_margin

earth_margin <- stats::qt(0.975, df = nrow(test)-1)*sd(earth_rmse)/sqrt(nrow(test))
earth_lower95 <- mean(earth_rmse) - earth_margin
earth_upper95 <- mean(earth_rmse) + earth_margin

linear_margin <- stats::qt(0.975, df = nrow(test)-1)*sd(linear_rmse)/sqrt(nrow(test))
linear_lower95 <- mean(linear_rmse) - linear_margin
linear_upper95 <- mean(linear_rmse) + linear_margin

randForest_margin <- stats::qt(0.975, df = nrow(test)-1)*sd(randForest_rmse)/sqrt(nrow(test))
randForest_lower95 <- mean(randForest_rmse) - randForest_margin
randForest_upper95 <- mean(randForest_rmse) + randForest_margin

svm_margin <- stats::qt(0.975, df = nrow(test)-1)*sd(svm_rmse)/sqrt(nrow(test))
svm_lower95 <- mean(svm_rmse) - svm_margin
svm_upper95 <- mean(svm_rmse) + svm_margin

tree_margin <- stats::qt(0.975, df = nrow(test)-1)*sd(tree_rmse)/sqrt(nrow(test))
tree_lower95 <- mean(tree_rmse) - tree_margin
tree_upper95 <- mean(tree_rmse) + tree_margin
```

Now we can show with integrity the mean value of the target column (medv), and the lower and upper 95% confidence estimates:

```{r Summary data frame with mean, lower and upper 95% confidence intervals}

summary_results <- 
  data.frame('Model' = c('bagging', 'bayesrnn', 'earth', 'linear', 'randForest', 'svm', 'tree'), 'Mean_results' = c(mean(bagging_rmse), mean(bayesrnn_rmse), mean(earth_rmse), mean(linear_rmse), mean(randForest_rmse), mean(svm_rmse), mean(tree_rmse)), 'Lower95_CI' = c(bagging_lower95, bayesrnn_lower95, earth_lower95, linear_lower95, randForest_lower95, svm_lower95, tree_lower95), 'Upper95_CI' = c(bagging_upper95, bayesrnn_upper95, earth_upper95, linear_upper95, randForest_upper95, svm_upper95, tree_upper95))

summary_results
```

This is only a few lines of code, well organized, that are returning results which are much stronger than a typical analysis. This analysis includes:

-   A range of models

-   Random resampling

-   95% confidence intervals around the root mean squared error

These steps produce results which predict in the real world with a higher degree of accuracy than an individual model run one time.

------------------------------------------------------------------------

## How to set up summary plots

### Predicted vs actual

One of the main ways you can demonstrate value to your customers is to show predicted vs actual information. We will start with plots, then generate summary tables.

```{r predicted vs actual plots}
bagging_plot <- data.frame(x = predict(object = bagging, newdata = test), y = test$y)
ggplot(bagging_plot, mapping = aes(x = x, y = y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Bagging predicted vs actual', x = 'Actual', y = 'Predicted')

bayesrnn_plot <- data.frame(x = predict(object = bayesRNN, newdata = test), y = test$y)
ggplot(bayesrnn_plot, mapping = aes(x = x, y = y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Bayesrnn predicted vs actual', x = 'Actual', y = 'Predicted')

earth_plot <- data.frame(x = predict(object = earth, newdata = test), y = test$y)
ggplot(earth_plot, mapping = aes(x = predict(object = earth, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Earth predicted vs actual', x = 'Actual', y = 'Predicted')

linear_plot <- data.frame(x = predict(object = linear, newdata = test), y = test$y)
ggplot(linear_plot, mapping = aes(x = predict(object = linear, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Linear predicted vs actual', x = 'Actual', y = 'Predicted')

randForest_plot <- data.frame(x = predict(object = randForest$best.model, newdata = test), y = test$y)
ggplot(randForest_plot, mapping = aes(x = predict(object = randForest$best.model, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Random Forest predicted vs actual', x = 'Actual', y = 'Predicted')

svm_plot <- data.frame(x = predict(object = svm$best.model, newdata = test), y = test$y)
ggplot(randForest_plot, mapping = aes(x = predict(object = svm$best.model, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Support Vector Machines predicted vs actual', x = 'Actual', y = 'Predicted')

tree_plot <- data.frame(x = predict(object = tree, newdata = test), y = test$y)
ggplot(randForest_plot, mapping = aes(x = predict(object = tree, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Linear predicted vs actual', x = 'Actual', y = 'Predicted')

```

## How to set up summary reports

1.  Set up the data
2.  Use the reactable package to make excellent tables which easily can be sorted, filtered, and searched.

```{r Set up summary tables}
library(reactable)
summary_table <- data.frame(
  'Model name' = c('Bagging', 'BayesRNN', 'Earth', 'Linear', 'Random Forest', 'Support Vector Machines', 'Trees'),
  'Mean RMSE' = c(mean(bagging_rmse), mean(bayesrnn_rmse), mean(earth_rmse), mean(linear_rmse), mean(randForest_rmse), mean(svm_rmse), mean(tree_rmse)),
  'Lower 95 CI' = c(bagging_lower95, bayesrnn_lower95, earth_lower95, linear_lower95, randForest_lower95, svm_lower95, tree_lower95),
  'Upper 95 CI' = c(bagging_upper95, bayesrnn_upper95, earth_upper95, linear_upper95, randForest_upper95, svm_upper95, tree_upper95)
  )

summary_table <- reactable::reactable(data = summary_table, sortable = TRUE, resizable = TRUE, filterable = TRUE, searchable = TRUE, showPageSizeOptions = TRUE)

summary_table
```

------------------------------------------------------------------------

Check your skills: We are going to repeat the analysis with a new data set, and use virtually everything we built to solve the Boston Housing data set.

::: {.callout-note appearance="minimal"}
The vast majority of what we need has already been built, tested, and verified to work for the Boston Housing data set. Now we are going to test the code with a totally new data set.
:::

## Intermediate data set: Numeric data, exploratory data analysis, and strings

Now we are going to add strings to our data set. The process will be extremely similar to what we just did with the Boston Housing data set. The new data set is a well known data set, commonly called the Insurance data set.

```{r Add strings to our numeric models}
df <- read.csv('https://raw.githubusercontent.com/InfiniteCuriosity/EnsemblesData/refs/heads/main/Insurance1.csv', stringsAsFactors = TRUE)
head(df)
str(df)
```

Now change the target column name to 'y':

```{r For the Insurance data set, change the target column name to y and test}
library(tidyverse)
df <- df %>% relocate(charges, .after = last_col())
df <- df %>% dplyr::rename(y = charges)
head(df)
```

As we can see the Insurance data set has seven columns. Three of the columns are imported as factors: sex, smoker, and region. We are now able to use them directly when we make our models.

<h4>Steps to building individual models in a consistent way:</h4>

We are going to follow the same steps we used for the Boston Housing data set, except that several of the columns are factors in the Insurance data set.

<h4>What does the customer actually want?</h4>

This is an insurance data set, the customer wants to model the charges (the seventh column). Specifically, they want to model the charges in a way that minimizes the charges they will pay, and meet their business obligations.

This is a very typical set of requirements from customers: Given the data from the customer, what is the best action the customer can take? Our modeling and recommendations will have a direct impact on their business decisions.

<h4>Before we run any code, let's ALWAYS take a look at the data to see what it suggests</h4>

In this section we are going to do an Exploratory Data Analysis of the Insurance data set. This was first published in 1961 by John Tukey. His definition states:

> "Procedures for analyzing data, techniques for interpreting the results of such procedures, ways of planning the gathering of data to make its analysis easier, more precise or more accurate, and all the machinery and results of (mathematical) statistics which apply to analyzing data." (add source)

```{r Boxplots of the numeric values in the Insurance data set}

df1 <- df %>% purrr::keep(is.numeric)
library(tidyverse)
df1 %>%
  tidyr::gather(key = "var", value = "value") %>%
  ggplot2::ggplot(aes(x = "", y = value)) +
  ggplot2::geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  ggplot2::facet_wrap(~var, scales = "free") +
  ggplot2::labs(title = "Boxplots of the numeric data")
```

```{r Histograms of the numeric values in the Insurance data set}

library(tidyverse)
ggplot2::ggplot(tidyr::gather(df1, cols, value), aes(x = value)) +
  ggplot2::geom_histogram(bins = round(nrow(df1))) +
  ggplot2::facet_wrap(. ~ cols, scales = "free") +
  ggplot2::labs(title = "Histograms of each numeric column.")
```

```{r Predictor (charges) vs each column}

df1 %>%
  tidyr::gather(-y, key = "var", value = "value") %>%
  ggplot2::ggplot(aes(x = value, y = y)) +
  ggplot2::geom_point() +
  ggplot2::facet_wrap(~ var, scales = "free") +
  ggplot2::theme_bw()+
  ggplot2::labs(title = "y (predictor variable) vs target variables")
```

Several results stand out from the exploratory data analysis, and we would do well to include them in our final report:

1.  The minimal charges to up as age goes up

2.  The maximum charges goes down as the number of children goes down

3.  Average and maximum charges for smokers are much higher than for non-smokers

<h4>Setting up models</h4>

We are able to use almost all of our code from our Boston housing example in the new insurance example. We will change the name of the variable to Y, initialize all of the model values, resample five times, and return summary results to the user.

```{r analysis for the Insurance data set, borrowing from our code for Boston Housing}

df$sex <- as.numeric(df$sex)
df$smoker <- as.numeric(df$smoker)
df$region <- as.numeric(df$smoker)

# Change the target variable ('Charges') to 'y'

names(df)[7] = 'y'

bagging_predicted <- 0
bagging_rmse <- 0
bayesRNN_rmse <- 0
earth_rmse <- 0
linear_rmse <- 0
randForest_rmse <- 0
svm_rmse <- 0
tree_rmse <- 0

for (i in 1:5) {
  sample <- sample(c(TRUE, FALSE), nrow(df), replace=TRUE, prob=c(0.6,0.4))
train  <- df[sample, ]
test   <- df[!sample, ]

bagging <- ipred::bagging(y ~ ., data = train)
bayesRNN <- brnn::brnn(x = as.matrix(train), y = train$y)
earth <- earth::earth(y ~ ., data = train)
linear <- lm(y ~ ., data = train)
randForest <- e1071::tune.randomForest(y ~ ., data = train)
svm <- e1071::tune.svm(y ~ ., data = train)
tree <- tree::tree(y ~ ., data = train)

bagging_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = bagging, newdata = test))
bagging_predicted[i] <- predict(object = bagging, newdata = test)

bayesRNN_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = bayesRNN, newdata = test))
bayesRNN_predicted[i] = predict(object = bayesRNN, newdata = test)

earth_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = earth, newdata = test))
earth_predicted[i] = predict(object = earth, newdata = test)

linear_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = linear, newdata = test))
randForest_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = randForest$best.model, newdata = test))
svm_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = svm$best.model, newdata = test))
tree_rmse[i] <- Metrics::rmse(actual = test$y, predicted = predict(object = tree, newdata = test))

results <- data.frame('Name' = c('bagging', 'bayesRNN', 'Earth', 'Linear', 'RandomForest', 'SVM', 'Treet'), 'RMSE' = c(mean(bagging_rmse), mean(bayesRNN_rmse), mean(earth_rmse), mean(linear_rmse), mean(randForest_rmse), mean(svm_rmse), mean(tree_rmse)))
results <- results %>% arrange(RMSE)
}
results
```

**Summary plots (predicted vs actual) for the Insurance data set**

```{r Summary plots for the Insurance data set}

bagging_plot <- data.frame(x = predict(object = bagging, newdata = test), y = test$y)
ggplot(bagging_plot, mapping = aes(x = x, y = y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Bagging predicted vs actual', x = 'Actual', y = 'Predicted')

bayesrnn_plot <- data.frame(x = predict(object = bayesRNN, newdata = test), y = test$y)
ggplot(bayesrnn_plot, mapping = aes(x = x, y = y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Bayesrnn predicted vs actual', x = 'Actual', y = 'Predicted')

earth_plot <- data.frame(x = predict(object = earth, newdata = test), y = test$y)
ggplot(earth_plot, mapping = aes(x = predict(object = earth, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Earth predicted vs actual', x = 'Actual', y = 'Predicted')

linear_plot <- data.frame(x = predict(object = linear, newdata = test), y = test$y)
ggplot(linear_plot, mapping = aes(x = predict(object = linear, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Linear predicted vs actual', x = 'Actual', y = 'Predicted')

randForest_plot <- data.frame(x = predict(object = randForest$best.model, newdata = test), y = test$y)
ggplot(randForest_plot, mapping = aes(x = predict(object = randForest$best.model, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Random Forest predicted vs actual', x = 'Actual', y = 'Predicted')

svm_plot <- data.frame(x = predict(object = svm$best.model, newdata = test), y = test$y)
ggplot(randForest_plot, mapping = aes(x = predict(object = svm$best.model, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Support Vector Machines predicted vs actual', x = 'Actual', y = 'Predicted')

tree_plot <- data.frame(x = predict(object = tree, newdata = test), y = test$y)
ggplot(randForest_plot, mapping = aes(x = predict(object = tree, newdata = test), y = test$y)) +
  ggplot2::geom_point() +
  ggplot2::labs(title = 'Linear predicted vs actual', x = 'Actual', y = 'Predicted')

```

**Summary plots (histograms of residuals, predicted vs residuals)**

```{r Histograms of residuals by model for the Insurance data set}

bagging_residuals <- data.frame('predicted' = c(bagging$y), 'residuals' = c(bagging$y - test$y))
ggplot(bagging_residuals, aes(x = residuals)) +
  geom_histogram(bins = round(nrow(df)/100)) +
  ggtitle('Histogram of bagging model residuals for the Insurance data set, each bar = 100 rows') +
  geom_vline(xintercept = 0, color = "red")

bayesRNN_residuals <- data.frame('predicted' = predict(object = bayesRNN, newdata = test), 'residuals' = c(predicted - test$y))
ggplot(bayesrnn_residuals, aes(x = residuals)) +
  geom_histogram(bins = round(nrow(df)/100)) +
  ggtitle('Histogram of bayesrnn model residuals for the Insurance data set, each bar = 100 rows') +
  geom_vline(xintercept = 0, color = "red")
```

<h4>The huge advantage is you've got a method you can use with a wide range of numeric data sets, without having to start from scratch every time</h4>

<h4>Lab 1: Introduction to individual numeric models</h4>

1.  Do an analysis of the New Boston Housing data set. The data is posted on the book's github site, and can be found here:

    <https://raw.githubusercontent.com/InfiniteCuriosity/Hands_On_Ensemble_Methods_With_R/refs/heads/main/Data/New_Boston_Housing.csv>

    Note the new data has strings (names) in the data, and the medv column has been moved. It is the 5th column in the New_Boston_Housing.csv data set. Also, the New_Boston_Housing data set includes corrected values for medv, the sixth column is named CMEDV, so that is the target column. You may drop the medv column from New_Boston_Housing.csv, as it has been corrected.

2.  Here is how to get started with New_Boston_Housing.csv:

    ```{r Importing New_Boston_Housing.csv}

    df <- read.csv('https://raw.githubusercontent.com/InfiniteCuriosity/Hands_On_Ensemble_Methods_With_R/refs/heads/main/Data/New_Boston_Housing.csv')
    df <- df[, -c(5)] # drops the 5th column, medv, because it has incorrect data
    head(df)
    str(df)
    ```

Practice building statistical models for the CMEDV column. Find the model with the lowest root mean squared error, and compare your results with the results for the original Boston Housing data set.

Include at least one new model in your analysis. Here are five possible choices (and there are many more options available):

```{r More models to test for the New_Boston_Housing data set}

install.packages('arm', repos = "http://cran.us.r-project.org") # for the Bayesglm package

install.packages('earth', repos = "http://cran.us.r-project.org") # Installs the earth package

install.packages('nnet', repos = "http://cran.us.r-project.org") # Installs the neuralnet package

install.packages('rpart', repos = "http://cran.us.r-project.org") # Installs the rpart package

install.packages('e1071', repos = "http://cran.us.r-project.org") # Installs the e1071 package, has several modeling methods available
```

```{r Modeling methods with the new packages}

bayesglm_train_fit <- arm::bayesglm(y ~ ., data = train, family = gaussian(link = "identity"))

earth_train_fit <- earth::earth(x = train[, 1:ncol(train) - 1], y = train$y)

neuralnet_train_fit <- nnet::nnet(train$y ~ ., data = train, size = 0, linout = TRUE, skip = TRUE)

rpart_train_fit <- rpart::rpart(train$y ~ ., data = train)

svm_train_fit <- e1071::tune.svm(x = train, y = train$y, data = train)
```

The original article was published in 1978 was focused on the pollution in the area. The pollution is measured in the NOX feature, the tenth column. Build statistical models to measure NOX. Resample 25 times, and report the root mean squared error for each of the models.

Produce diagnostic plots for each of the models: Predicted vs actual, histogram of residuals, histograms, and Q-Q plot.

#### Help for you building these models:

Look in the Appendix in this book. There are completed examples of reports written in Quarto that show the entire process with the New_Boston_Housing data set, Wine Quality Color, AutoMPG (miles per gallon), and Facebook Metrics.

**Post on social media**

You may use any social media you prefer, such as TikTok, YouTube, BlueSky, Instagram, or any other social media platform of your choice. Post on social media about your work in this chapter. Use at least three hashtags, and tag at least two other people in your post.
